\subsection*{Architecture Differences}
\addcontentsline{toc}{section}{Architecture Differences}

In order to understand the GPGPU hardware architecture differences, one must first understand the hardware and the reasoning behind it's design. The GPU, in particular, was designed around multimedia functions for a computer that operated on the timescale of a human eye, rather than the timescale of the CPU itself. The human eye works in terms of hundredths of a second, rather than pico- or nano- seconds. Also, the GPU works with huge amounts of data at any given time. Each pixel on the screen has several different attributes, for example, the RGBa common layout describes each pixel position in the vector in terms of Red, Green, Blue, and Alpha channel attributes. When applied to modern 1080p screens, it is not hard to comprehend how much data the GPU must process. \cite{gpucomputing}

For example, on a typical 1080p screen with a standard $1920 X 1080$ resolution:
\begin{center}
\begin{gather}
1920 X 1080 = 2,073,600 \text{ px} \notag \\
2,073,600 \text{ px} * 4 \text{(r/g/b/a)} =  8,294,400 \notag
\end{gather}
\end{center}

That means, for a 1080p screen, the GPGPU must perform at least 8,294,400 calculations for each screen refresh at 60 fps (assuming the standard 60hz refresh rate)! That requires a ton of computation power! 

Luckily, as stated above, it must work in slower time-slices than the CPU (60 refreshes per second). This allows for a much deeper (in the hundreds to thousands deep) pipeline. The latency of each individual instruction does not mater as much as the aggregate throughput to the screen. Therefore, to understand the design choices around the GPGPU's architecture, one must understand it was designed around the theory that "throughput is more important than latency." \cite{gpucomputing}

Where as the typical CPU pipeline works by breaking up each task into steps then feeding the output of each step into the input of the next step and donating all of the CPUs resources to that time slice, the GPU works differently. Instead of dividing the processor an the work in terms of \textit{time}, the GPU divides the work in terms of \textit{space}. Each stage in the GPU pipeline represents a different space of the GPU. This works extremely well because the GPU can exploit the data parallelism from within each \textit{space}. \cite{gpucomputing} Exploiting that level of parallelism allows the GPU to compute multiple data elements at the same time (SIMD). 

\subsection*{How does a GPGPU process data?}
\addcontentsline{toc}{section}{How does a GPGPU process data?}

As discussed in the previous section, GPGPUs utilize the Single Instruction Multiple Data (SIMD) architecture. In this architecture, all executing threads must move through the computation in lock-step. Threads cannot deviate from the execution path. Ideally threads would execute without branches or control flow operations that would effect their route through the GPGPU. The efficiency of a GPGPU (SIMD efficiency) is defined as "the fraction of scalar threads that execute together in lockstep per cycle" \cite{memcached}

GPGPU architecture varies from a general purpose CPU architecture in two main ways, Control Flow and Memory Access. These all end up as challenges when incorporating the GPGPU into the CPU to create the heterogeneous architecture. 

\subsection*{Control Flow}
\addcontentsline{toc}{subsection}{Control Flow}

Recall that in a SIMD (Single Instruction Multiple Data) architecture, all concurrently executing threads (also known as wavefronts) must move through the GPGPU in lock-step. Therefore, the ideal situation for these wavefronts (groups of threads representing divided work) would be to move through the code with no conditional branches allowing unabated forward progress through the program and correctly executing instructions for each work item (data set) in parallel. \cite{memcached}

However, this is almost never the case. Modern mature programs almost always include control flow branches in the code. With that being said, it is entirely possible for threads in the wavefront to encounter data items that cause the code to make the jump, while the rest of the threads do not. Predictably, this causes problems when threads are expected to execute in lock step. This is known as "branch divergence" and is currently solved with the addition of a hardware component that monitors the running threads. This SIMT stack monitor watches threads executing and disables threads making an uncharacteristic jump. It tracks the current program counter and the RPC (re-convergence program counter), which is the closest point in the code at which the inactive and active threads will rejoin. \cite{memcached, gpuarch}

This does not cause the threads that made the jump to miss out on any execution. As soon as the active threads finish executing their code the converge and the inactive threads continue execution. The main difference here is that on a CPU, if threads on executing on different cores diverge, they can all continue execution on their own paths. Each core can be at a vastly different place in the program but still actively executing code. However, the GPGPU runs like a train, and must sequentially walk through the entire program, so divergent threads, rather than executing in different places, inactivate for the points in the code they do not execute. \cite{memcached}

\subsection*{Memory Access}
\addcontentsline{toc}{subsection}{Memory Access}

The differences in memory access can be divided into two sections: main memory access, and cache management. 

\subsubsection*{Main Memory Management}
\addcontentsline{toc}{subsubsection}{Main Memory Management}

Another major considerations for programs running on a GPGPU is memory access. A GPGPU can run thousands of threads simultaneously. If the currently executed instruction on the GPGPU is a memory access instruction (like a load or store), then each thread will generate a memory access request. If this memory request happens to be for a different address then the GPGPU will have to handle these requests individually. However, if the program was designed with the GPGPU in mind and high amounts of data locality amongst threads in the wavefront is achieved, then the GPGPU can optimize this request itself and consolidate the memory requests into a single request. \cite{memcached, tlpcache}

This is not something normally thought about when writing general programs for a CPU. Since most consumer CPU's can run between 8 and 12 threads at once, if they generate memory requests the overall performance impact is not a big deal. However, the scale of the GPGPU can make this into a big problem. \cite{memcached}

\subsubsection*{Cache Management}
\addcontentsline{toc}{subsubsection}{Cache Management}

Another point to mention regarding memory access is the cache management. While both the GPGPU and the CPU can maintain their own caches, they access data in the cache in very different ways. This makes combining them very difficult. GPGPUs hide memory access latency by utilizing a massive number of threads. This requires a huge cache to manage the constant context switching on the GPGPU. This causes the GPGPU to hit the cache much more frequently than the CPU will. Assuming a general LRU replacement policy, the cache will clearly favor the GPGPU. \cite{tlpcache}

\subsection*{Making them work together}
\addcontentsline{toc}{subsection}{Making them work together}

It seems like combining a CPU and GPGPU is a large challenge because of the fundamental differences between the two pieces of hardware. However, because of the massive gains in performance that a GPGPU has over the equivalent CPU SIMD instruction set (Intel SEE, for example), there are huge gains to be had from successfully combining them. 

The key in making the heterogenous architecture successful is remembering that an ideal heterogenous architecture will support the GPGPU core as an individual processing unit much like an additional CPU core. Therefore, the control flow differences become a lesser problem because the CPU will be waiting for the GPGPU to finish, but does not need to be tied to its lock step computation. This will allow the GPGPU to act much like a thread would under a standard processor, dispatching work to it, and waiting for the result. 












